User Guide for Corrfunc

### 2015-01-28 paper/users-guide.tex
Added code to the python script to read in the ascii file using pandas (if pandas is installed)
Almost complete users-guide.

\documentclass[12pt,titlepage]{article}

\usepackage{amsmath}
\usepackage{aas_macros}
\usepackage{rotating}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{color}
\usepackage{listings}
\usepackage{caption}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{dirtree}
\usepackage{adjustbox}
\usepackage{bytefield}
\usepackage{chngcntr}

%% stdsection is the old section -> no new page
%% section produces a new page before a new section
\let\stdsection\section
\renewcommand\section{\newpage\stdsection}

%% Some of the short-cuts I use
\newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
\newcommand{\xir}{\ensuremath{{\xi(r)}}\xspace}
\newcommand{\wprp}{\ensuremath{{w_p(r_p)}}\xspace}
\newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
\newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
\newcommand{\clang}{{\texttt{clang}}\xspace}
\newcommand{\icc}{{\texttt{icc}}\xspace}
\newcommand{\gcc}{{\texttt{gcc}}\xspace}

@@

%% Code paper - so there are lots of code listings
\lstset{
  language=C,                % choose the language of the code
  basicstyle=\ttfamily,      % sets font
  numbers=none,                   % where to put the line-numbers
  numberstyle=\small, 
  stepnumber=1,                   % the step between two line-numbers.
  numbersep=10pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  tabsize=2,                      % sets default tabsize to 2 spaces
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=true,         % sets if automatic breaks should only happen at whitespace
}
%% Change the listings counter to show Section number + listings number (rather than just the listings number)
\AtBeginDocument{\counterwithin{lstlisting}{section}}

@@

\section{Introduction}
Correlation functions are a statistical measure of a density field and are widely used in large-scale structure formation. Generally, 
the measurements are done {\em once}\footnote{which is why I have not bothered with releasing the codes to measure correlation 
functions on data} on survey data and compared with model predictions in a Monte-Carlo Markov Chain. As such, the 
correlation functions have to be measured repeatedly during an MCMC. The codes presented here are meant to cover the typical scenarios 
of measuring correlation functions in theory-land. The primary consideration in writing these codes is speed\footnotemark -- the codes 
presented here should outperform any other CPU based correlation functions codes by a wide-margin.

\footnotetext{The secondary consideration was maintainability and ease of use for others. I have versions of these codes that are 
even faster but are potentially gibberish for anyone other than me!}

@@

The directory structure for the code looks like this:
\dirtree{%
.1 corrfunc.
.2 paper.
.2 xi\_theory.
.3 benchmarks\DotFill
  \begin{minipage}[t]{8cm}
    IDL scripts to run benchmarks{.} 
  \end{minipage}.
.3 bin\DotFill 
  \begin{minipage}[t]{8cm}
    Will be created to copy executable files when you run `make install'{.} 
  \end{minipage}.
.3 examples\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for example C bindings using the static libraries{.} 
  \end{minipage}.
.3 include\DotFill 
  \begin{minipage}[t]{8cm}
    Header files for static libraries{.} 
  \end{minipage}.
.3 io\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for reading in data{.} 
  \end{minipage}.
.3 lib\DotFill 
  \begin{minipage}[t]{8cm}
    Will be created to copy static libraries and python library after you run `make'{.} 
  \end{minipage}.
.3 python\_bindings\DotFill 
  \begin{minipage}[t]{8cm}
    Source files to generate python bindings{.} 
  \end{minipage}.
.3 tests\DotFill 
  \begin{minipage}[t]{8cm}
    Correct outputs for tests{.}
  \end{minipage}.
.4 data\DotFill 
  \begin{minipage}[t]{8cm}
    Mock galaxy catalogs for tests{.} 
  \end{minipage}.
.3 utils\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for creating 3-D grid and helper routines{.} 
  \end{minipage}.
.3 wp\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for $w_p(r_p)${.} 
  \end{minipage}.
.3 xi\_of\_r\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for $\xi(r)${.} 
  \end{minipage}.
.3 xi\_rp\_pi\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for $\xi(r_p,\pi)${.} 
  \end{minipage}.
}

@@

\begin{table}
\begin{center}
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{ccccM{4in}} 
\toprule
\multicolumn{1}{c}{\textbf{Option Type}}   &
\multicolumn{1}{c}{\textbf{Option Name}}   &
\multicolumn{1}{c}{\textbf{Default State}} &
\multicolumn{1}{c}{\textbf{Requires}}      &
\multicolumn{1}{c}{\textbf{Notes}}      \\
\midrule
                 & PERIODIC            & Enabled  & None            & Enables periodic boundary conditions. \\
\textbf{Science} & OUTPUT\_RPAVG       & Disabled & DOUBLE\_PREC    & Outputs the average pair-separation in each bin. \xir and \wprp can be slower by more than $2\times$, \xirppi is less affected. \\
\midrule
                 & DOUBLE\_PREC        & Disabled & None                               & Computations are done using double precision. Slower and requires more RAM. \\
\textbf{Code}    & USE\_AVX            & Enabled  & CPU and compiler with AVX support  & CPUs later than 2011 have AVX support. Code will run much faster with this option. \\
                 & USE\_OMP            & Enabled  & OpenMP capable compiler            & Since \clang does not support OpenMP yet, \texttt{common.mk} will stop compilation with \clang when this flag is enabled. \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{center}
\caption{\footnotesize List of compilations options, what the options mean and their dependencies for the codes. }
\label{table:options}
\end{table}
Depending on your Science use-case and the cpu/compiler, you will want to set the different options. Once you set those options, you should set the 
C compiler, CC (available options are \icc, \gcc, \clang). Once you have set the compiler, installing should be as simple as typing `make' and 
`make install' in the \texttt{xi\_theory} directory. All the libraries are intentionally chosen to be \texttt{static} libraries just to avoid 
any path conflicts. However, on MAC OSX, you may have to do more to get the library to work -- so I have outlined some of the scenarios in 
Section~\ref{section:mac}. 
\subsection{Linux}
If the installation went well, you should have an executable called \texttt{run\_correlations} in the \texttt{examples} directory. Type \texttt{./run\_correlations} 
in the \texttt{examples} directory and you should see the code in action. The C source file \texttt{run\_correlations.c} also serves as an example to 
use the \xir, \xirppi and \wprp libraries in C.

@@

\section{Running the Codes}
\subsection{Input Files}
The codes currently can handle these types of input data files:
\begin{itemize}
\item \texttt{ascii} -- White-space separated columns, format code is `a'.
\item \texttt{csv}   -- Comma-separated values, format code is `c'.
\item \texttt{fast-food} -- Fast-food, fortran binary format, format code is `f'.
The fast-food file format is described in detail in Section~\ref{section:fastfood}.
\end{itemize}
For the \texttt{ascii} and \texttt{csv} files, the code reads in the first three columns as the co-moving \texttt{X/Y/Z} arrays. Note, that more columns 
can be present but the code will ignore those columns.

@@

\flushleft The \texttt{fast-food} file consists of a header:
\begin{lstlisting}[language=C,numbers=none]
int idat[5];
float fdat[9];
float znow;
\end{lstlisting}
For the purposes of these correlation function codes, the only useful quantity is \texttt{idat[1]} which contains \texttt{N} -- the number of particles 
in the data file.

\vspace{0.2in}
%% \begin{enumerate}
%% \item
\begin{bytefield}{28}
  \bitheader{0-27} \\
  \bitbox{4}{20} & \bitbox{4}{Int} & \bitbox{4}{\textcolor{red}{\texttt{N}}} \bitbox{12}{3 Integers}& \bitbox{4}{20}\\
\end{bytefield}\\
%% \item
%% \begin{bytefield}{44}
%%   \bitheader{0-43} \\
%%  \bitbox{4}{36} & \bitbox{36}{9 Floats} & \bitbox{4}{36} \\
%% \end{bytefield}
%% %% \item
%% \begin{bytefield}{12}
%%   \bitheader{0-11} \\
%%   \bitbox{4}{4} & \bitbox{4}{Float} & \bitbox{4}{4}\\
%% \end{bytefield}
%% \end{enumerate}
The next 56 bytes contain two other fields (10 floats + 4 padding bytes of 4 bytes each) and their corresponding paddings. Since these bytes do not contain any data 
that are useful in the context of these codes we will skip over the contents of these bytes. In \texttt{io.c}, I actually \texttt{fseek} over these bytes.

\flushleft After this header, the actual \texttt{X/Y/Z} values are stored. The first 4 bytes after the header contains \texttt{N*4, for float precision} or \texttt{N*8, for 
double precision} where \texttt{N=idat[1]}, is the number of particles in the file. After all the \texttt{X} values there will be another 4 bytes containing 
\texttt{N*4} or \texttt{N*8}. {\em Note, that even when the \texttt{X/Y/Z} arrays are written out in double-precision, the padding is still 4 bytes.}
\begin{minipage}[t]{\textwidth}
\centerline{\underline{\bfseries Byte-structure of the X/Y/Z arrays in a fast-food file}}
\vspace{0.2in}
\begin{minipage}[t]{0.5\textwidth}
\centering
\textbf{floats}\\[3ex]
\begin{bytefield}{4}
  \bitheader{0-3} \\
  \bitbox{4}{\texttt{N*4}} \\
  \wordbox{1}{X[0]} \\
  \wordbox{1}{X[1]} \\
  \wordbox[]{1}{$\vdots$} \\[1ex]
  \wordbox{1}{X[N-1]} \\
  \bitbox{4}{\texttt{N*4}}
\end{bytefield} 
\end{minipage}%
\begin{minipage}[t]{0.5\textwidth}
\centering
\textbf{doubles}\\[3ex]
\begin{bytefield}{8}
  \bitheader{0-7} \\
  \bitbox{4}{\texttt{N*8}} \\
  \wordbox{1}{X[0]} \\
  \wordbox{1}{X[1]} \\
  \wordbox[]{1}{$\vdots$} \\[1ex]
  \wordbox{1}{X[N-1]} \\
  \bitbox{4}{\texttt{N*8}}
\end{bytefield}
\end{minipage}
\end{minipage}
\subsection{Specifying the radial bins}\label{section:bins}
The codes were intentionaly designed to read in a set of (somewhat) arbitrary\footnote{I assume bins are non-overlapping} set of bin specifications from a file. This way, you can specify disjoint bin-edges as well as 
use $0.0$ as a bin edge (which would be impossible if $\log$ bins are assumed). The bins are to be specified in a (white-space separated) text file in this manner:
\begin{verbatim}
r_low[0]  r_high[0]
r_low[1]  r_high[1]
.
.
.
r_low[nbins-1]  r_high[nbins-1]
\end{verbatim}
where, \texttt{r\_low[i]} and \texttt{r\_high[i]} are the left and right edges of the \texttt{i}'th bin respectively. The text files should contains as many lines as the number of 
bins desired. The \texttt{logbins} executable can be used to create such a text file containing $\log$ bins. The syntax for running \texttt{logbins} is:
\begin{lstlisting}[language=sh,numbers=none]
./logbins <rmin> <rmax> <nbins>   > <text filename>
\end{lstlisting}
An example of such a file with radial bins is the file \texttt{bins} in the \texttt{tests} directory.

{\em Note, all of the three codes print the correlation function to \texttt{stdout} -- so be sure to redirect \texttt{stdout} to an output file. }

\subsection{Running \texorpdfstring{\xir}{xi(r)}}
To run the correlation function in 3-D, you will need to run the executable \texttt{DD} from either the \texttt{bin} or \texttt{xi\_of\_r} directory. The inputs 
to \texttt{DD} are:
\begin{itemize}
\item \texttt{file1}   -- the file name for the first file.
\item \texttt{format1} -- the file format for the first file.
\item \texttt{file2}   -- the file name for the second file.
\item \texttt{format2} -- the file format for the second file.
\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
\end{itemize}
Thus, the code can compute an auto-correlation (when \texttt{file1,format1} and \texttt{file2,format2} are identical) or a cross-correlation (when \texttt{file1} and \texttt{file2} are different). Out of the box, 
some sample \texttt{DD} call directory) can be:
\begin{itemize}
\item auto-correlation --  {\scriptsize \texttt{./DD  ../tests/data/gals\_Mr19.ff f ../tests/data/gals\_Mr19.ff f ../tests/bins 4 > Mr19\_output.DD}}
\item cross-correlation -- {\scriptsize \texttt{./DD ../tests/data/cmassmock\_Zspace.ff f ../tests/data/random\_Zspace.ff f ../tests/bins 4 > cmass\_output.DR}}
\end{itemize}
The output (printed to stdout) has \texttt{nbins} columns \texttt{<Npairs> <rpavg> <rmin> <rmax>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
Makefile option OUTPUT\_RPAVG has been enabled. Here, each line of the output represents the \texttt{i}'th radial bin.

\subsection{Running \texorpdfstring{\xirppi}{xi(rp,pi)}}
To run the correlation function in 2-D for \xirppi, you will need to run the executable \texttt{DDrppi} from either the \texttt{bin} or \texttt{xi\_rp\_pi} directory. The inputs 
to \texttt{DDrppi} are:
\begin{itemize}
\item \texttt{file1}   -- the file name for the first file.
\item \texttt{format1} -- the file format for the first file.
\item \texttt{file2}   -- the file name for the second file.
\item \texttt{format2} -- the file format for the second file.
\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
\item \texttt{pimax}   -- the maximum distance to consider in the $\pi$ direction. The code uses \texttt{Z} axis as the $\pi$ (line-of-sight) direction. 
\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
\end{itemize}
Thus, the code can compute an auto-correlation (when \texttt{file1,format1} and \texttt{file2,format2} are identical) or a cross-correlation (when \texttt{file1} and \texttt{file2} are different). Out of the box, 
some sample \texttt{DDrppi} call directory) can be:
\begin{itemize}
\item auto-correlation --  {\scriptsize \texttt{./DDrppi  ../tests/data/gals\_Mr19.ff f ../tests/data/gals\_Mr19.ff f ../tests/bins 40.0 4 > Mr19\_output\_rppi.DD}}
\item cross-correlation -- {\scriptsize \texttt{./DDrppi ../tests/data/cmassmock\_Zspace.ff f ../tests/data/random\_Zspace.ff f ../tests/bins 80.0 4 > cmass\_output\_rppi.DR}}
\end{itemize}
The output (printed to stdout) has \texttt{nbins} columns \texttt{<Npairs> <rpavg> <log(rmin)> <pi\_upper>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
Makefile option OUTPUT\_RPAVG has been enabled. The code bins in 1 Mpc/h in the $\pi$ direction by default. Therefore, 
the total number of bins in the output file will be $nbins \times \pi_{\rm max}$, where each radial bin is further broken into $\pi_{\rm max}$ bins along the 
$\pi$ direction.

@@

>>>>

\documentclass[12pt,titlepage,justified]{article}

\usepackage{amsmath}
\usepackage{aas_macros}
\usepackage{rotating}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{color}
\usepackage{listings}
\usepackage{caption}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{dirtree}
\usepackage{adjustbox}
\usepackage{bytefield}
\usepackage{chngcntr}
\usepackage{alltt}

%% stdsection is the old section -> no new page
%% section produces a new page before a new section
\let\stdsection\section
%% \renewcommand\section{\newpage\stdsection}

%% Some of the short-cuts I use
\newcommand{\rmax}{\ensuremath{{r_{max}}}\xspace}
\newcommand{\xir}{\ensuremath{{\xi(r)}}\xspace}
\newcommand{\wprp}{\ensuremath{{w_p(r_p)}}\xspace}
\newcommand{\xirppi}{\ensuremath{{\xi(r_p,\pi)}}\xspace}
\newcommand{\pimax}{\ensuremath{{\pi_{\rm max}}}\xspace}
\newcommand{\hMpc}{\ensuremath{{h^{-1}Mpc}\xspace}}
\newcommand{\todo}[1]{\marginpar{TODO}{\color{red}#1}}
\newcommand{\clang}{{\texttt{clang}}\xspace}
\newcommand{\icc}{{\texttt{icc}}\xspace}
\newcommand{\gcc}{{\texttt{gcc}}\xspace}

@@

%% Code paper - so there are lots of code listings
\lstset{
  language=C,                % choose the language of the code
  basicstyle=\scriptsize\ttfamily,      % sets font
  numbers=none,                   % where to put the line-numbers
%%   numberstyle=\small, 
%%   stepnumber=1,                   % the step between two line-numbers.
%%   numbersep=10pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  tabsize=2,                      % sets default tabsize to 2 spaces
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=true,         % sets if automatic breaks should only happen at whitespace
  captionpos=t,  
  xleftmargin=17pt,
  framexleftmargin=17pt,
  framexrightmargin=17pt,
  framexbottommargin=5pt,
  framextopmargin=5pt,
  showstringspaces=false,
  frame=top,frame=bottom,
  escapeinside={(*@}{@*)},
}
%% Change the listings counter to show Section number + listings number (rather than just the listings number)
\AtBeginDocument{\counterwithin{lstlisting}{section}}

\DeclareCaptionFormat{listing}{\rule{\dimexpr\textwidth+17pt\relax}{0.4pt}\par\vskip1pt#1#2#3}
\captionsetup[lstlisting]{format=listing,singlelinecheck=false, margin=0pt, font={sf},labelsep=space,labelfont=bf}

%%
\renewcommand{\thetable}{\Roman{table}}
\captionsetup[table]{margin=0pt,labelsep=period,labelfont=bf,textfont={it}}
\captionsetup[figure]{margin=0pt,labelsep=period,labelfont=bf,textfont={it}}

@@

\section{Introduction}
Correlation functions are a statistical measure of a density field and are widely used in large-scale structure formation. Generally, 
the measurements are done {\em once}\footnote{which is why I have not bothered with releasing the codes to measure correlation 
functions on data} on survey data and compared with model predictions in a Monte-Carlo Markov Chain. As such, the 
correlation functions have to be measured repeatedly during an MCMC. The codes presented here are meant to cover the typical scenarios 
of measuring correlation functions in theory-land. The primary consideration in writing these codes is speed\footnotemark -- the codes 
presented here should outperform any other CPU based correlation functions codes by a wide margin.

Cache locality and hand-written AVX intrinsics are the reasons why the code is very fast. However, that also means that the code is 
not very portable. I have tried my best to ensure that the codes work on Linux and MAC OSX. If it does not work for you, particularly 
if you are on a reasonable Linux install, please \href{mailto:manodeep@gmail.com}{email me}. The API for the codes can be considered 
frozen; I will not change the API without changing the MAJOR release version.

The paper associated with the codes is being (slowly) written by me. I hope to submit the paper to Astronomy \& Computing and release 
the codes simultaneously. Once I submit the paper to arXiv, there will be an actual paper arXiv link \href{http://adsabs.harvard.edu/abs/2014arXiv1405.5832L}{here}.

\footnotetext{The secondary consideration was maintainability and ease of use for others. I have versions of these codes that are 
even faster but are much harder to modify/maintain by any one other than me !}

@@

The directory structure for the code looks like this:
\dirtree{%
.1 corrfunc.
.2 paper.
.2 xi\_theory.
.3 benchmarks\DotFill
  \begin{minipage}[t]{8cm}
    IDL scripts to run benchmarks{.} 
  \end{minipage}.
.3 bin\DotFill 
  \begin{minipage}[t]{8cm}
    Will be created to copy executable files when you run `make install'{.} 
  \end{minipage}.
.3 examples\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for example C bindings using the static libraries{.} 
  \end{minipage}.
.3 include\DotFill 
  \begin{minipage}[t]{8cm}
    Header files for static libraries{.} 
  \end{minipage}.
.3 io\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for reading in data{.} 
  \end{minipage}.
.3 lib\DotFill 
  \begin{minipage}[t]{8cm}
    Will be created to copy static libraries and python library after you run `make'{.} 
  \end{minipage}.
.3 python\_bindings\DotFill 
  \begin{minipage}[t]{8cm}
    Source files to generate python bindings{.} 
  \end{minipage}.
.3 tests\DotFill 
  \begin{minipage}[t]{8cm}
    Correct outputs for tests{.}
  \end{minipage}.
.4 data\DotFill 
  \begin{minipage}[t]{8cm}
    Mock galaxy catalogs for tests{.} 
  \end{minipage}.
.3 utils\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for creating 3-D grid and helper routines{.} 
  \end{minipage}.
.3 xi\_of\_r\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for $\xi(r)${.} 
  \end{minipage}.
.3 xi\_rp\_pi\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for $\xi(r_p,\pi)${.} 
  \end{minipage}.
.3 wp\DotFill 
  \begin{minipage}[t]{8cm}
    Source files for $w_p(r_p)${.} 
  \end{minipage}.
}

@@

\begin{table}
\centering
\caption{\footnotesize List of compilations options, what the options mean and their dependencies for the codes. }
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{ccccM{4in}} 
\toprule
\multicolumn{1}{c}{\textbf{Option Type}}   &
\multicolumn{1}{c}{\textbf{Option Name}}   &
\multicolumn{1}{c}{\textbf{Default State}} &
\multicolumn{1}{c}{\textbf{Requires}}      &
\multicolumn{1}{c}{\textbf{Notes}}      \\
\midrule
                 & PERIODIC            & Enabled  & None            & Enables periodic boundary conditions. \\
\textbf{Science} & OUTPUT\_RPAVG       & Disabled & DOUBLE\_PREC    & Outputs the average pair-separation in each bin. \xir and \wprp can be slower by more than $2\times$, \xirppi is less affected. \\
\midrule
                 & DOUBLE\_PREC        & Disabled & None                               & Computations are done using double precision. Slower and requires more RAM. \\
\textbf{Code}    & USE\_AVX            & Enabled  & CPU and compiler with AVX support  & CPUs later than 2011 have AVX support. Code will run much faster with this option. \\
                 & USE\_OMP            & Enabled  & OpenMP capable compiler            & Since \clang does not support OpenMP yet, \texttt{common.mk} will stop compilation with \clang when this flag is enabled. \\
\bottomrule
\end{tabular}
\end{adjustbox}
\label{table:options}
\end{table}
Depending on your Science use-case and the cpu/compiler, you will want to set the different options. Once you set those options, you should set the 
C compiler, CC (available options are \icc, \gcc, \clang). Once you have set the compiler, installing should be as simple as typing `make' and 
`make install' in the \texttt{xi\_theory} directory. All the libraries are intentionally chosen to be \texttt{static} libraries just to avoid 
any path conflicts. However, on MAC OSX, you may have to do more to get the library to work -- so I have outlined some of the scenarios in 
Section~\ref{section:mac}. 
\subsection{Linux}
If the installation went well, you should have an executable called \texttt{run\_correlations} in the \texttt{examples} directory. Type \texttt{./run\_correlations} 
in the \texttt{examples} directory and you should see the code in action. The C source file \texttt{run\_correlations.c} also serves as an example to 
use the \xir, \xirppi and \wprp libraries in C.

@@

\subsection{Running the tests}
If installation went fine, then run \texttt{make tests} to run a suite of tests. If any of the tests fail, then please \href{mailto:manodeep@gmail.com}{email me}. I have 
never seen the tests fail unless I made some coding error while modifying the source. 
Once the tests pass successfully, type \texttt{make install} to install the binaries into the \texttt{bin} directory.

\stdsection{Running the Codes}
The codes should run straight out of the box. Note, the \xir and \xirppi can compute both auto and cross-correlations with and without PERIODIC boundary conditions whereas the \wprp code 
only computes the auto-correlation with PERIODIC boundary conditions. Also, since both the \xir and \xirppi codes compute cross-correlations, pairs are {\em double-counted} in both 
\xir and \xirppi codes. You could, in theory, make the auto-correlation bit faster by only computing unique pairs (as is done in \wprp, see Section~\ref{section:pair_counting_wp} 
for details). However, since the total run-time will dominated by the cross-correlation (the number of randoms is typically an order of magnitude larger than the number 
of data points), I have not implemented those optimizations for the auto-correlation calculations. 
\subsection{Input Files}
The codes currently can handle these types of input data files:
\begin{itemize}
\item \texttt{ascii} -- White-space separated columns, format code is `a'.
\item \texttt{csv}   -- Comma-separated values, format code is `c'.
\item \texttt{fast-food} -- Fast-food, fortran binary format, format code is `f'.
The fast-food file format is described in detail in Section~\ref{section:fastfood}.
\end{itemize}
For the \texttt{ascii} and \texttt{csv} files, the code reads in the first three columns as the co-moving \texttt{X/Y/Z} arrays. Note, that more columns 
can be present but the code will ignore those columns.

@@

\flushleft The \texttt{fast-food} file consists of a header:
\begin{lstlisting}[language=C,numbers=none,caption={The header format for \texttt{fast-food} files}]
int idat[5];
float fdat[9];
float znow;
\end{lstlisting}
For the purposes of these correlation function codes, the only useful quantity is \texttt{idat[1]} which contains \texttt{N} -- the number of particles 
in the data file.

\vspace{0.2in}
\begin{bytefield}{28}
  \bitheader{0-27} \\
  \bitbox{4}{20} & \bitbox{4}{Int} & \bitbox{4}{\textcolor{red}{\texttt{N}}} \bitbox{12}{3 Integers}& \bitbox{4}{20}\\
\end{bytefield}\\
The next 56 bytes contain two other fields (10 floats + 4 padding bytes of 4 bytes each) and their corresponding paddings. Since these bytes do not contain any data 
that are useful in the context of these codes we will skip over the contents of these bytes. In \texttt{io.c}, I actually \texttt{fseek} over these bytes.

After this header, the actual \texttt{X/Y/Z} values are stored. The first 4 bytes after the header contains \texttt{N*4} for float precision or \texttt{N*8} for 
double precision where \texttt{N=idat[1]}, is the number of particles in the file. After all the \texttt{X} values there will be another 4 bytes containing 
\texttt{N*4} or \texttt{N*8}. {\em Note, that even when the \texttt{X/Y/Z} arrays are written out in double-precision, the padding is still 4 bytes.} The 
blocks for \texttt{Y/Z} follow after the \texttt{X} block.

\vspace{0.2in}
\begin{minipage}[t]{\textwidth}
\centerline{\underline{\bfseries Byte-structure of the X/Y/Z arrays in a fast-food file}}
\vspace{0.2in}
\begin{minipage}[t]{0.5\textwidth}
\centering
\textbf{floats}\\[3ex]
\begin{bytefield}{4}
  \bitheader{0-3} \\
  \bitbox{4}{\texttt{N*4}} \\
  \wordbox{1}{X[0]} \\
  \wordbox{1}{X[1]} \\
  \wordbox[]{1}{$\vdots$} \\[1ex]
  \wordbox{1}{X[N-1]} \\
  \bitbox{4}{\texttt{N*4}}
\end{bytefield} 
\end{minipage}%
\begin{minipage}[t]{0.5\textwidth}
\centering
\textbf{doubles}\\[3ex]
\begin{bytefield}{8}
  \bitheader{0-7} \\
  \bitbox{4}{\texttt{N*8}} \\
  \wordbox{1}{X[0]} \\
  \wordbox{1}{X[1]} \\
  \wordbox[]{1}{$\vdots$} \\[1ex]
  \wordbox{1}{X[N-1]} \\
  \bitbox{4}{\texttt{N*8}}
\end{bytefield}
\end{minipage}
\end{minipage}
\subsection{Specifying the radial bins}\label{section:bins}
The codes were intentionaly designed to read in a set of (somewhat) arbitrary\footnote{I assume bins are non-overlapping} set of bin specifications from a file. This way, you can specify disjoint bin-edges as well as 
use $0.0$ as a bin edge (which would be impossible if $\log$ bins are assumed). The bins are to be specified in a (white-space separated) text file in this manner:
\begin{alltt}
r\_low[0]  r\_high[0]
r\_low[1]  r\_high[1]
\vdots
r\_low[nbins-1]  r\_high[nbins-1]
\end{alltt}
where, \texttt{r\_low[i]} and \texttt{r\_high[i]} are the left and right edges of the \texttt{i}'th bin respectively. The text files should contains as many lines as the number of 
bins desired. The \texttt{logbins} executable can be used to create such a text file containing $\log$ bins. The syntax for running \texttt{logbins} is:\\
\vspace{0.2in}
\texttt{./logbins <rmin> <rmax> <nbins>   > <text filename>}
\vspace{0.2in}
\\An example of such a file with radial bins is the file \texttt{bins} in the \texttt{tests} directory. 
{\em Note, all of the three codes print the correlation function to \texttt{stdout} -- so be sure to redirect \texttt{stdout} to an output file. }

\subsection{Running \texorpdfstring{\xir}{xi(r)}}
To run the correlation function in 3-D, you will need to run the executable \texttt{DD} from either the \texttt{bin} or \texttt{xi\_of\_r} directory. Note, that 
\xir double-counts the pairs (as does \xirppi). The inputs to \texttt{DD} are:
\begin{itemize}
\item \texttt{file1}   -- the file name for the first file.
\item \texttt{format1} -- the file format for the first file.
\item \texttt{file2}   -- the file name for the second file.
\item \texttt{format2} -- the file format for the second file.
\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
\end{itemize}
Thus, the code can compute an auto-correlation (when \texttt{file1,format1} and \texttt{file2,format2} are identical) or a cross-correlation (when \texttt{file1} and \texttt{file2} are different). Out of the box, 
some sample \texttt{DD} call directory) can be:
\begin{itemize}
\item auto-correlation --  {\scriptsize \texttt{./DD  ../tests/data/gals\_Mr19.ff f ../tests/data/gals\_Mr19.ff f ../tests/bins 4 > Mr19\_output.DD}}
\item cross-correlation -- {\scriptsize \texttt{./DD ../tests/data/cmassmock\_Zspace.ff f ../tests/data/random\_Zspace.ff f ../tests/bins 4 > cmass\_output.DR}}
\end{itemize}
The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<Npairs> <rpavg> <rmin> <rmax>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
Makefile option OUTPUT\_RPAVG has been enabled. Here, each line of the output represents the \texttt{i}'th radial bin.

\subsection{Running \texorpdfstring{\xirppi}{xi(rp,pi)}}
To run the correlation function in 2-D for \xirppi, you will need to run the executable \texttt{DDrppi} from either the \texttt{bin} or \texttt{xi\_rp\_pi} directory. Note, that 
\xirppi double-counts the pairs (as does \xir).The inputs 
to \texttt{DDrppi} are:
\begin{itemize}
\item \texttt{file1}   -- the file name for the first file.
\item \texttt{format1} -- the file format for the first file.
\item \texttt{file2}   -- the file name for the second file.
\item \texttt{format2} -- the file format for the second file.
\item \texttt{binfile} -- the file name for the file containing the bins (see Section~\ref{section:bins})
\item \texttt{pimax}   -- the maximum distance to consider in the $\pi$ direction. The code uses \texttt{Z} axis as the $\pi$ (line-of-sight) direction. 
\item \texttt{nthreads} -- the number of OpenMP threads to use (only required when the Makefile option USE\_OMP is enabled)
\end{itemize}
Thus, the code can compute an auto-correlation (when \texttt{file1,format1} and \texttt{file2,format2} are identical) or a cross-correlation (when \texttt{file1} and \texttt{file2} are different). Out of the box, 
some sample \texttt{DDrppi} call directory) can be:
\begin{itemize}
\item auto-correlation --  {\scriptsize \texttt{./DDrppi  ../tests/data/gals\_Mr19.ff f ../tests/data/gals\_Mr19.ff f ../tests/bins 40.0 4 > Mr19\_output\_rppi.DD}}
\item cross-correlation -- {\scriptsize \texttt{./DDrppi ../tests/data/cmassmock\_Zspace.ff f ../tests/data/random\_Zspace.ff f ../tests/bins 80.0 4 > cmass\_output\_rppi.DR}}
\end{itemize}
The output (printed to stdout) has \texttt{nbins} rows; each row contains the columns \texttt{<Npairs> <rpavg> <log(rmin)> <pi\_upper>}, where the \texttt{<rpavg>} column contains \texttt{0.0} unless the 
Makefile option OUTPUT\_RPAVG has been enabled. The code bins in 1 Mpc/h in the $\pi$ direction by default. Therefore, 
the total number of bins in the output file will be \texttt{nbins$\times$\pimax}, where each radial bin is further broken into \pimax bins along the 
$\pi$ direction.

@@

\section{License}
The code has been released under the MIT License. 
\begin{verbatim}
Permission is hereby granted, free of charge, to any person 
obtaining a copy of this software and associated documentation 
files (the ``Software''), to deal in the Software without 
restriction, including without limitation the rights to use, 
copy, modify, merge, publish, distribute, sublicense, and/or
sell copies of the Software, and to permit persons to whom 
the Software is furnished to do so, subject to the following 
conditions:

The above copyright notice and this permission notice shall 
be included in all copies or substantial portions of the 
Software.
